### 2019-09-09 manuscript/manuscript.tex
upated castro figure

  \caption{{\bf A}: The convergence correlation, averaged across replicate
    pairs, through time. Each line represents the convergence correlation
    $\mathrm{cor}(\Delta p_{s}, \Delta p_{s})$ from a starting reference
    generation $s$ to a later time $t$, which varies along the x-axis; each
    line corresponds to a row of the temporal convergence correlation matrix
    depicted to the right.  {\bf B}: The convergence covariance between
    individual pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf
    C}:  The convergence covariance between individual pairs of replicates in
\parencite{Castro2019-uk} data, for the two selection lines  (LS1 and LS2) and
the control (Ctrl).}

@@

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to changing epistatic
relationships among alleles or  recombination breaking up selective alleles.
Such reversals of selective dynamics could be occurring at other timepoints but
the signal of a change in the direction of selection at particular loci may be
washed out when we calculate our genome-wide average temporal covariances.  To
address this limitation, we calculated the distribution of the temporal
covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance of each tile will be noisy, due to sampling and genetic drift, and
the neutral distribution of the covariance is complicated due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). \vb{To address
this}, we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the   allele frequency changes
per-genomic window. This destroys the systematic covariances created by linked
selection \vb{and} creates a sampling distribution of the covariances
spuriously created by neutral genetic drift while preserving the complex
dependencies between adjacent loci created by linkage disequilibrium (see
Supplementary Material Section \ref{supp:empirical-null} for more details on
this approach).  This empirical neutral null distribution is conservative in
the sense that the variances of the covariances are wider than expected under
drift alone as they include the effect of selection on the allele frequency
change within a time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend \vince{GC: you had ??? here, not sure I see the issue.} (Supplementary
Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that
the windowed covariances not only decay towards zero, but in fact become
negative through time, consistent with many regions in the genome having had a
reversed fitness effect at later timepoints.

@@

Second, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response of
selection of about 5 standard deviations over the course of twenty generations
\parencite{Castro2019-uk}. This study includes two independent selection lines,
Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
Consequently, this selection experiment offers a useful control to test our
between-replicate covariances: we expect to see positive between-replicate
covariance in the comparison between the two Longshanks selection lines, but
not between the two pairwise comparisons between the control line and the two
Longshanks lines. We find that this is the case (Figure \ref{fig:figure-2} C),
with the two Longshanks comparisons to the control line not being significantly
different from zero, while the comparison between the two Longshanks line is
statistically significantly different from zero (CIs [0.013, 0.037]).

\gc{Paragraph about large effect loci.}
%We find that for covariances 2 timepoints apart ($k=2
%\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
%[1.29, 1.32]), consistent with linked selection creating positive covariance in
%allele frequency change. Under constant directional selection, we expect these
%positive covariances between increasingly distant timepoints to decay to zero
%as additive genetic variance is exhausted by selection and linkage
%disequilibrium decays. Instead we find that for covariances 4 timepoints apart
%($k=4 \times 10$ generations), the left tail of negative covariances is
%inflated by a factor of 1.19 (95\% CI [1.17, 1.21]).

>>>>

  \caption{{\bf A}: The convergence correlation, averaged across replicate
    pairs, through time. Each line represents the convergence correlation
    $\mathrm{cor}(\Delta p_{s}, \Delta p_{s})$ from a starting reference
    generation $s$ to a later time $t$, which varies along the x-axis; each
    line corresponds to a row of the temporal convergence correlation matrix
    depicted to the right.  {\bf B}: The convergence covariance between
    individual pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf
    C}:  The convergence covariance between individual pairs of replicates in
    \parencite{Castro2019-uk} data, for the two selection lines  (LS1 and LS2)
    and the control (Ctrl); gray CIs are those using the complete dataset, blue
    CIs exclude chromosomes 5 and 10 which harbor the two regions
    \textcite{Castro2019-uk} found to have signals of parallel selection between
    LS1 and LS2.}

@@

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to changing epistatic
relationships among alleles or  recombination breaking up selective alleles.
Such reversals of selective dynamics could be occurring at other timepoints but
the signal of a change in the direction of selection at particular loci may be
washed out when we calculate our genome-wide average temporal covariances.  To
address this limitation, we calculated the distribution of the temporal
covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance of each tile will be noisy, due to sampling and genetic drift, and
the neutral distribution of the covariance is complicated due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). \vb{To address
this}, we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the   allele frequency changes
per-genomic window. This destroys the systematic covariances created by linked
selection \vb{and} creates a sampling distribution of the covariances
spuriously created by neutral genetic drift while preserving the complex
dependencies between adjacent loci created by linkage disequilibrium (see
Supplementary Material Section \ref{supp:empirical-null} for more details on
this approach). This empirical neutral null distribution is conservative in
the sense that the variances of the covariances are wider than expected under
drift alone as they include the effect of selection on the allele frequency
change within a time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
striking pattern that the windowed covariances not only decay towards zero, but
in fact become negative through time, consistent with many regions in the
genome having had a reversed fitness effect at later timepoints.

% TODO \vince{GC: you had ??? here, not sure I see the issue.}

@@

Second, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response of
selection of about 5 standard deviations over the course of twenty generations
\parencite{Castro2019-uk}. This study includes two independent selection lines,
Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
Consequently, this selection experiment offers a useful control to test our
between-replicate covariances: we expect to see positive between-replicate
covariance in the comparison between the two Longshanks selection lines, but
not between the two pairwise comparisons between the control line and the two
Longshanks lines. We find that this is the case (gray confidence intervals in
Figure \ref{fig:figure-2} C), with the two Longshanks comparisons to the
control line not being significantly different from zero, while the comparison
between the two Longshanks line is statistically significantly different from
zero (CIs [0.013, 0.037]).

A major finding in the Longshanks study were two major-effect loci that showed
parallel frequency shifts between the two selection lines: a region harboring
the gene Nkx3-2 known to be involved in limb development, and another region
harboring six other candidate genes. We were curious to what extent our
genome-wide covariances were being driven by these two outlier large-effect
loci, so we excluded them from the analysis. Since we do not know the extent to
which linkage disequilibrium around these large-effect loci affects neighboring
loci, we took the conservative precaution of excluding the entire chromosomes
these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
covariances. We find excluding these large effect loci has little impact on the
confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
indicating that we these across-replicate covariances are indeed driven by a
polygenic signal.

%We find that for covariances 2 timepoints apart ($k=2
%\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
%[1.29, 1.32]), consistent with linked selection creating positive covariance in
%allele frequency change. Under constant directional selection, we expect these
%positive covariances between increasingly distant timepoints to decay to zero
%as additive genetic variance is exhausted by selection and linkage
%disequilibrium decays. Instead we find that for covariances 4 timepoints apart
%($k=4 \times 10$ generations), the left tail of negative covariances is
%inflated by a factor of 1.19 (95\% CI [1.17, 1.21]).
